---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()

```


#Reformat Catch Data

* remove  "all" column
* create species column
    - move from wide to long
* general QA

[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1)

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)
```

Remove the `all` and `notesRegCode` columns using `select`
Ctrl+shift+m: pipe operator shortcut %>%


```{r}
catch_long <- catch_original %>% 
    select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)%>% #select(-All, -notesRegCode)
    gather(key = "species", value = "catch", -Region, -Year)


head(catch_long)

```
```{r}
catch_wide <- catch_long %>% 
    spread(key = Region, value = catch)

head(catch_wide)
```

* rename catch to catch_thousands
* change catch column to numeric
* create a new catch column in units num. of fish `#`

```{r}
catch_clean <- catch_long %>% 
    rename(catch_thousands = catch) %>%
    mutate(catch_thousands = ifelse(catch_thousands %in% c("I"), 1, catch_thousands)) %>% 
    mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
    mutate(catch = catch_thousands *1000) %>% 
    select(-catch_thousands)

head(catch_clean)

```

```{r, eval = F, echo = F}
#Look for where as.numeric failed 
test <- as.numeric(catch_long$catch)
i <- which(is.na(test) == TRUE)
catch_long[i,]

```

##Split - Apply - Combine

* calculate mean catch by species

#num_obs = n()

```{r}
year_mean <- catch_clean %>% 
    group_by(Year, species) %>% 
    summarise(year_mean = mean(catch),num_obs = length(which(is.na(catch)))) %>%   
  #summarise(year_mean = mean(catch), num_obs =length(which(is.na(catch)))) %>%  
    arrange(-Year, desc(species)) %>%
    filter(Year <= 1990 & Year >=1960)

#head(species_mean)
year_mean

```




#Join the Region Definitions

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea) #%>% 
    #rename(Region,code)

head(region_defs)

```

```{r}
catch_joined <- left_join(catch_clean, region_defs, by = c("Region" = "code"))

#catch_joined <- catch_clean %>%
#    left_join(region_defs)

head(catch_joined)

```


#Misc. Functions

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

```


```{r}
dates_split <- dates_df %>%
    separate(date, into = c("month", "day", "year"), sep ="/", remove = F)

head(dates_split)

```

```{r}
dates_unite <- dates_split %>%
    unite(col = date_iso, year, month, day, sep ="-")

head(dates_unite)

```

```{r}
library(stringr)

month <- c("5", "05", "03")

str_pad(month, 2, side = "left", pad ="0")

```

```{r}
write.csv(catch_clean, "catch_clean.csv", row.names=F)

```





