---
title: "Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(leaflet)
library(DT)

```



# Reformat the data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r, echo = F}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = FALSE)
```

* calculate annual escapment by species and region

* filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)

```{r, echo = F}
annual_esc <- esc %>%
    separate(sampleDate, into = c("year", "month", "day"), sep ="-") %>% 
    mutate(year = as.numeric(year)) %>% 
    group_by(year, Species, SASAP.Region) %>% 
    summarise(count = sum(DailyCount)) %>% 
    filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink")) 
  #coerce year to numeric year
  #ungroup() %>% 
 

head(annual_esc)
```



# Genberate some static plots

```{r, echo = F}
#ggplot(annual_esc) + geom_col(mapping = aes(x = Species, y = count), fill = "light yellow")

ggplot(annual_esc, mapping = aes(x = Species, y = count, fill = SASAP.Region)) + #scale_fill_brewer(palette = "Dark2") +
geom_col()

```

```{r, echo = F}
my_theme <- theme_test() +
  theme(legend.position = "bottom", legend.title = element_blank(), plot.title = element_text(hjust=0.5, color="black"))

```



```{r, echo = F}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = count, color = Species)) + geom_line()+
  geom_point() +
  ylab("Escapement (num of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  theme_test() +
  my_theme


```
```{r, fig.width = 7, fig.height = 8, echo = F}
p <- ggplot(annual_esc, aes(x = year, y = count, color = Species)) + geom_line() +
  geom_line() +
  geom_point() +
  scale_y_continuous(label = comma) +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
  my_theme
p

ggsave("figures/regional_escapement.png",
       plot = p,
       width = 7,
       height = 8,
       units = "in")

```

 
# Generate an interactive map
Map the sampling locations for escapement

<br/>
<br/>
 
* find distinct lat/lon/location combinations

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>%  #use - to drop any column not be included
  drop_na() #drop any rows with NA


```

```{r, echo = F}
datatable(locations)

```

```{r}
leaflet(locations) %>% 
    addTiles() %>% 
    addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~ Location)

```

For dyrple/tidyr reference see my [data tidying page](https://cwang2019.github.io/arctic-training/data-tiding.html)

<br/>
<br/>
<br/>